---
title: Product Insights Through Trelliscope
thumbnail: "/img/trelliscope_img.png"
author: Sunita Kenner
date: '2017-06-30'
slug: product-insights-through-trelliscope
categories:
  - Business Science
  - Product Insights
  - Visualizations
tags:
  - trelliscopejs
  - R
  - multiple displays
---

**Note: The trelliscope plot displays with blogdown would not have been possible without [Yihui Xie's](https://yihui.name) lightning fast help. Not only did he implement the fix (in an afternoon), but [documented/provided 2 options](https://bookdown.org/yihui/blogdown/dep-path.html) to fixing the trelliscope display issue. Thank you so much Yihui.**

One of the impressive presentations at [Jan 2017 RStudio Conference](https://www.rstudio.com/conference/) was Ryan Hafen's [trelliscopejs R package in action: Visualization in the Tidyverse](http://slides.com/hafen/trelliscopejs#/). "TrelliscopeJS is a framework for creative interactive displays of small multiples", and this is incredibly powerful when providing insights on product lines to decision makers.  

```{r, echo = FALSE, results = 'hide', message= FALSE, warning= FALSE, error=FALSE}
devtools::install_github("hafen/trelliscopejs")
library(trelliscopejs)
library(tidyverse)
library(rbokeh)
#install.packages("ggplot2")


```

```{r, echo = FALSE, warning = FALSE}
Product_df<- readRDS("~/Documents/Blog trellis/Product_Data/Product_df.rds")

Product_df_1 <- Product_df %>% 
  filter(Region != " Latin America " & Region != " APAC ") %>% 
  select(Products, Unit.Cost, Amount, Qty, Ship.Country, Year, Month, region, industry)
```

User will be able to use the `Sort`, `Filter`, `Labels`, and `Grid` tabs to gain further insights from the data. A powerful way to dig deeper into data without having to produce several seperate plots. The display can be viewed full screen using the toggle @ right-hand corner.

Using `nest()` in `tidyverse`, `rbokeh` and `trelliscope`

```{r, warning = FALSE, message = FALSE}

Trellis_Data1 <- Product_df_1 %>%
  group_by(Products, Month, region) %>% 
  mutate(Unit.Cost = mean(Unit.Cost))

Trellis_Data2 <- Trellis_Data1 %>%
  group_by(Products, region) %>%
  mutate(industry_most = names(tail(sort(table(industry)), 1)))

 Trell_plot <- Trellis_Data2 %>%
  dplyr::group_by(Products, Month, region, industry_most) %>%
  dplyr::summarise(Qty = sum(Qty), Amt = sum(Amount), ASP = Amt/Qty) %>%
  group_by(Products, region) %>%
  nest() %>%
  mutate(panel = map_plot(data,
    ~ figure(toolbar = NULL, xlim = c(4.5, 12.5), ylim = c(21.5, 101)) %>%
      ly_points(Month, ASP, hover = .x[1:3], data = .x)
  ))

Trell_plot %>% trelliscope(name = "ASP_Month", nrow =2, ncol = 4, width = 500,  path = blogdown::dep_path())

```

The one variable I was not seeing on the display to sort and filter by was `industry`.  Please see note from Ryan Hafen for an explanation as to why the variable cannot be used for display, and what can be done to address it. 

```{r, eval = FALSE}
Trellis_Data1 %>%
  group_by(Products, region) %>%
  summarise(uniq_industry = length(unique(industry)))

```

###Note from Ryan Hafen

> To add additional variables to a display to sort and filter on, those variables just need to be present in the data. If it's a numeric variable, a summary statistic (mean) will be computed. If it's categorical, like `industry`, then the variable can only be included if it is constant within each panel (each panel can have a set of scalar metrics to sort and filter on - these cannot be vectors). The `industry` variable cannot be used in the above display because it is not unique within each combination of `Products` and `region`. You could create a new variable that is the most prominent industry for each `Products` / `region` combination and then it would be automatically included.

```{r}
Trellis_Data2 <- Trellis_Data1 %>%
  group_by(Products, region) %>%
  mutate(industry_most = names(tail(sort(table(industry)), 1)))

Trellis_Data2
```

Using `qplot` and `facet_treplliscope`

```{r, eval = FALSE}

#Does not work because of ggplot2
install.packages("ggplot2")
library(ggplot2)
qplot(Month, Unit.Cost, data = Trellis_Data2) + 
  xlim(4.5, 12.5) + ylim(21.5, 101) + 
facet_trelliscope(~ Products + region, nrow = 2, ncol = 4, width = 500, path = blogdown::dep_path("rmarkdown_files/Unitcost_vs_Month_gg"))

```

There will be future posts on gaining powerful multiple insights from displays of small multiples.  

Thanks again for all your help [Yihui](https://yihui.name) & [Ryan](https://hafen.github.io/trelliscopejs/#trelliscope). Appreciate the time and effort you've spent in helping me resolve some of the issues I encountered while using the package.  
[More trelliscopejs](http://ryanhafen.com/blog/trelliscopejs). 
