---
title: From Powerpoint to ShinyDashboards & Highcharter
thumbnail: "/img/ppt_dashboards.png"
author: Sunita Kenner
date: '2017-11-15'
slug: from-powerpoint-to-shinydashboards
categories:
  - Competitive Brief
  - Shiny Dashboards
tags:
  - R
  - blogdown
  - shiny
  - Highcharter
---


One of my goals for 2017 was to build and provide presentations as interactive products distributed and accessible to the management team as a web app.  Using [ShinyDashboards](https://rstudio.github.io/shinydashboard/get_started.html) and [highcharter: R wrapper for Highcharts](http://jkunst.com/highcharter/) for visualization, here is a [starter example on shinyapps.io](https://busaanalytics.shinyapps.io/compBrief_Blogdown/) of one of the presentations that I would have normally created in Powerpoint. Also embedded in the post.

**Please note that Highcharts (http://highcharts.com) is a Highsoft product which is not free for commercial and Governmental use.**

<iframe width="900" height="900" scrolling="yes" frameborder="no"  src="//busaanalytics.shinyapps.io/compBrief_Blogdown/"> </iframe>


```{r, eval = FALSE, echo = FALSE}
# Used iframe to control the size of the app. 
knitr::include_app("https://busaanalytics.shinyapps.io/compBrief_Blogdown/", height = "900px")
#knitr::include_app("https://busaanalytics.shinyapps.io/compBrief_Blogdown/", height = "900px", width = "900px")
```

There are plenty of [ShinyDashboards Examples with source code](https://rstudio.github.io/shinydashboard/examples.html) and [Joshua Kunst](http://jkunst.com/) who is the author of Highcharter has done an excellent job of providing examples on how to replicate [highcharts demos](https://cran.r-project.org/web/packages/highcharter/vignettes/replicating-highcharts-demos.htmlauthor). 

Thank you [Joshua Kunst](http://jkunst.com/) for all your help/responses on the highchart/highcharter visualization questions I had. 

Here is an example of the code for the [column with drilldown chart](https://www.highcharts.com/demo/column-drilldown) on my dataset.

```{r, error= FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(highcharter)
```

```{r, echo = FALSE}

mkt_share <-
  readxl::read_excel("~/Documents/Blog trellis/Product_Data/blogdown_data.xlsx",
  sheet = "blogdown_mkt_Share data")
  all_mkt_share <-
  readxl::read_excel("~/Documents/Blog trellis/Product_Data/blogdown_data.xlsx",
  sheet = "Other")
  
  mkt_share <- mkt_share %>%
  filter(Year == 2016) %>%
  arrange(desc(`Annual Rev`))
  
  other_data <- all_mkt_share[-(1:19), ] %>%
  arrange(desc(`Annual Rev`))

```

View market share data

```{r}
mkt_share
other_data
```

```{r, echo = FALSE}
# highchart() %>% 
#  hc_add_series(Prop_survey_data, "pie", hcaes(name = Mode_of_trans, y = prop_of_students), name = "% of Trans", dataLabels = list(enabled = TRUE, format = '{point.name}: {point.y}%')) %>% 
#  hc_add_theme(hc_theme_ffx())
```

Prep the data for drill-down

```{r}
#First data set

first_mkt_share_data <- mkt_share %>%
  mutate(name = Vendor,
         y = `Annual Rev`,
         drilldown = tolower(name)) %>%
  select(name, y, drilldown) %>%
  list_parse()

#Second data set containing drill-down content
data_for_drilled_column <- other_data %>%
  mutate(name = Vendor,
         y = `Annual Rev`,
         drilldown = tolower(name)) %>%
  select(name, y) %>%
  list_parse2() %>%
  map(function(x) {
    x[[2]] <- as.numeric(x[[2]])
    x
})

# Plot using highcharter

hc <- highchart() %>%
  hc_chart(type = "column") %>%
  hc_title(text = "2016: WW Product Market Share $M") %>%
  hc_subtitle(text = "Click 'Other' to view 'Other' Vendors") %>%
  hc_xAxis(type = "category") %>%
  hc_legend(enabled = FALSE) %>%
  hc_plotOptions(series = list(boderWidth = 0, dataLabels = list(enabled = TRUE))) %>%
  hc_add_series(name = "Mkt Share",colorByPoint = TRUE, data = first_mkt_share_data) %>%
  hc_drilldown(allowPointDrilldown = TRUE,
  series = list(list(id = "other", data = data_for_drilled_column)))
  
hc

```

There are plenty of [ShinyDashboards Examples with source code](https://rstudio.github.io/shinydashboard/examples.html) and [Joshua Kunst](http://jkunst.com/) who is the author of Highcharter has done an excellent job of providing examples on how to replicate highcharts demos. 

Thank you [Joshua Kunst](http://jkunst.com/) for all your help on the visualizations.  
Other resources for highcharter visualizations that I found useful in creating the visuals:
*  [Danton Noriega's post](https://dantonnoriega.github.io/ultinomics.org/post/2017-04-05-highcharter-explainer.html)
*  [R documentation on highcharter package](https://www.rdocumentation.org/packages/highcharter/versions/0.5.0/topics/hc_drilldown)


