---
title: "Forecasting W/Prophet Vs Hyndman’s Forecast Package"
author: "Sunita Kenner"
date: '2017-05-01'
slug: forecasting-w-prophet-vs-hyndman-s-forecast-package
tags:
- R
- Prophet
- Forecast
- dygraph
categories:
- Forecasting
- Time Series
---

Having used [Rob Hyndman's forecasting package](http://robjhyndman.com) in R, I decided to give [Facebook's Prophet](https://facebookincubator.github.io/prophet/docs/quick_start.html) package a spin and do a comparison of the results.

```{r, echo = FALSE, message = FALSE, warning=FALSE}
#load required libraries

library(forecast)
library(tidyverse)
#library(dygraphs)
library(plotly)
library(lubridate)
library(prophet)

#devtools::install_github('hadley/ggplot2')

```


```{r, echo = FALSE, eval = FALSE, warning=FALSE}
#Load data sets:

daily_data_raw <- readRDS("~/Documents/Blog trellis/Product_Data/daily_data.rds")
monthly_data_raw <- readRDS("~/Documents/Blog trellis/Product_Data/monthlydata.rds")

data_manipulate <- function(df) {
  df %>% 
    filter(y >0) %>% 
    mutate(y = log(y), y_new = round(y*5)) %>%
    select(y_new, ds) %>%
    rename(y = y_new)
}

data_manipulate_month <- function(df) {
  df %>% 
    filter(y >0) %>% 
    mutate(y = round(y*25)) %>%
    select(y, ds)
}

daily_data <- data_manipulate(daily_data_raw)
saveRDS(daily_data, "~/Documents/Blog trellis/Product_Data/daily_data1.rds")
monthly_data <- data_manipulate_month(monthly_data_raw)
saveRDS(monthly_data, "~/Documents/Blog trellis/Product_Data/monthly_data1.rds")

```

```{r, echo = FALSE}
daily_data <- readRDS("~/Documents/Blog trellis/Product_Data/daily_data1.rds")
monthly_data <- readRDS("~/Documents/Blog trellis/Product_Data/monthly_data1.rds")
```

##Forecasts using Prophet

The Time-series data sets:
- Daily data set

```{r, echo = FALSE}
head(daily_data)
```

Prophet provides the ability to [forecast projected growth](https://facebookincubator.github.io/prophet/docs/forecasting_growth.html) based on market intel, by specifying the carrying capacity 'cap'; the forecast should reach the maximum/saturation at the set 'cap'. For this data set, I've set the cap at 7 units/day, but the forecaster can choose to set it at an increasing/decreasing sequence.  

```{r}

daily_Data_wgrowth <- daily_data %>% 
  mutate(cap = 7)
head(daily_Data_wgrowth)
```

Visualize the daily timeseries
```{r}
ggplotly(ggplot(daily_data, aes(x = ds, y = y)) + geom_line(color = "firebrick") + ggtitle("Daily Timeseries")) 
```

Functions calling function prophet that fits and returns the model, then predicts on the model object. More info available from [Prophet's page](https://facebookincubator.github.io/prophet/docs/installation.html) 

```{r}
#Function to apply prophet model to timeseries
prophet_apply <- function(timeseries, growthtype = NULL) {
  if(is.null(growthtype))
     prophet(timeseries)
    else
      prophet(timeseries, growth = growthtype)
}

#Create future data_frame and predict on the model object
prophet_fcst_apply <-
  function(m,
           length_time,
           cap_size = NULL) {
    if (is.null(cap_size))
    { future_df <- make_future_dataframe(m, periods = length_time)
    }
    else 
    { future_df <- make_future_dataframe(m, periods = length_time)
      future_df['cap'] = cap_size
    }
    
  fcst_df <- predict(m, future_df)
  }

```


Model and forecast with prophet:

```{r, warning= FALSE, message = FALSE}
model_daily <- prophet_apply(daily_data)
model_daily_wgrowth <- prophet_apply(daily_Data_wgrowth, 'logistic')

Fcst_daily <- prophet_fcst_apply(model_daily, 730)
Fcst_daily_wgrowth <- prophet_fcst_apply(model_daily_wgrowth, 730, 7)

```

View forecasted units with their upper/lower confidence interval 
```{r}

#Select forecast with lower/upper confidence intervals from the predicted object
#Specify time frame for selection
view_select_Fcst <- function(fcst_df, date_begin, date_end) {
  selected_Fcst <- fcst_df %>% 
    filter(ds >date_begin  & ds< date_end) %>% 
    select(ds, yhat, yhat_lower, yhat_upper)
}

tail(view_select_Fcst(Fcst_daily, "2017-04-01", "2018-04-01"))
```

Visualize the forecast by calling the generic plot function. I'm wrapping ggplotly(via [Plotly package](https://plot.ly/ggplot2/)) around the plot function for interactive web-based version.

Plots for forecasted daily data: 
```{r, warning=FALSE, message=FALSE}

#Plot function for daily fcst
plot_daily <- function(m, fcst, fnc) {
    fnc(m, fcst)
}

ggplotly(plot_daily(model_daily, Fcst_daily, plot))
plot_daily(model_daily, Fcst_daily, prophet_plot_components)

```

Sum forecast unit totals and forecast order dollars based on user specified price (in this case. price is set at $25).

```{r}
#summarise yearly forecast units and order dollars
summary_fcst_units_price <- function(df, price, fcst_num) {
  df %>% 
   dplyr::summarise(Fcst_qty = sum(df[[fcst_num]]), Fcst_Dlrs = Fcst_qty*price)
}

```

Daily data set:
```{r}
#Call summary_fcst_units_price function on the predicted daily numbers
sum_daily <- summary_fcst_units_price(view_select_Fcst(Fcst_daily, "2017-04-01", "2018-04-01"), 25, "yhat")
sum_daily
```

Plots for daily_data with growth

```{r, message = FALSE, warning= FALSE}
ggplotly(plot_daily(model_daily_wgrowth, Fcst_daily_wgrowth, plot))
plot_daily(model_daily_wgrowth, Fcst_daily_wgrowth, prophet_plot_components)

```

Daily_data with growth set:
```{r}
tail(view_select_Fcst(Fcst_daily_wgrowth, "2017-04-01", "2018-04-01"))
sum_daily_wgrowth <- summary_fcst_units_price(view_select_Fcst(Fcst_daily_wgrowth,  "2017-04-01", "2018-04-01"),25, "yhat")
sum_daily_wgrowth

```


## Forecasts using Forecast Package by Hyndman

Using Auto ETS:
```{r}

#Auto ETS forecast
daily_ts <- ts(daily_data[,1], start(2016,1,1), frequency = 7)
daily_auto_fcst <- forecast(daily_ts, h = 365)
#autoplot(daily_auto_fcst)
ggplotly(autoplot(daily_auto_fcst))

```

Sum total units and total order dollars based on price at $25
```{r}
sum_hyndman_ets <- summary_fcst_units_price(data.frame(daily_auto_fcst), 25, "Point.Forecast")
```

Using Auto Arima:
```{r}
#Auto Arima Forecast
fit_ar_daily<- auto.arima(daily_ts)
daily_auto_ar_fcst <- forecast(fit_ar_daily, h = 365)
#autoplot(daily_auto_fcst)
ggplotly(autoplot(daily_auto_fcst))

```

```{r}
sum_hyndman_ar <- summary_fcst_units_price(data.frame(daily_auto_ar_fcst), 25, "Point.Forecast")

```

Let's now compare the forecast results from Prophet and Forecast package by Hyndman
```{r, message = FALSE}
All_TS_fcst <- full_join(full_join(sum_daily, sum_daily_wgrowth), full_join(sum_hyndman_ets, sum_hyndman_ar))
rownames(All_TS_fcst) <- c("Fcst_daily_prophet", "Fcst_daily_wgrowth_prophet", "Fcst_daily_forecast_ETS", "Fcst_daily_forecast_ARIMA")
knitr::kable(All_TS_fcst, caption = "Forecasts from Prophet & Forecast Package")
```
