---
title: Sales Insights using Shiny Modules & Highcharter
thumbnail: "/img/shinymodules_hc.png"
author: Sunita Kenner
date: '2018-03-05'
slug: shiny-modules-product-line-insights
categories:
  - Business Science
  - Product Insights
  - Pricing
tags:
  - R
  - blogdown
  - Highcharter
---

To further my 2017 objective of delivering presentations or analysis' results via an interactive app, I attended RStudio's Shiny app training course at the [January 2018 RStudio Conference](https://www.rstudio.com/conference/).  

One of the business issues was to figure out a balanced distribution of the world wide and US sales force so as to maximize order dollars.  This required figuring out each Product Line's geographical strength and weaknesses in terms of `Order $$` and a further drill-down on each of the top performers within the product line.  Generating insights for each of the products and a summary per their Product Line meant reusing the same shiny app code for each of those products, for which [shiny modules](https://shiny.rstudio.com/articles/modules.html) provide an efficient approach.  

Here is a [Minimal example on shinyapps.io](https://busaanalytics.shinyapps.io/Module_hc_blogdown_app/). Also embedded in the post.

**Please note that Highcharts (http://highcharts.com) is a Highsoft product which is not free for commercial and Governmental use.**

<iframe width="900" height="900" scrolling="yes" frameborder="no"  src="//busaanalytics.shinyapps.io/Module_hc_blogdown_app/"> </iframe>


Many thanks to [Winston Chang](https://github.com/wch) who sacrificed his breaks at the RStudio Shiny App training course to patiently answer and help me resolve some of the issues I had with the development of this biz dashboard.  Thanks to the RStudio community member @nutterb (Benjamin) for introducing me to scoped verbs in dplyr which resolved the issue posted [here](https://community.rstudio.com/t/issue-making-dplyr-reactive-variables-work-in-shiny-module/3694).

RStudio has quite a few tutorials on [Shiny Modules](https://www.rstudio.com/resources/webinars/understanding-shiny-modules/). Joe Cheng's article on Modularizing [Shiny App Code](https://shiny.rstudio.com/articles/modules.html).  Other: 

Here is the server code that calls the module

```{r, eval = FALSE}

source("alt_prod_blog.R")

ui <- dashboardPage(
  # Application title ---------------------------------------------------------
  dashboardHeader(title = "Business Insights", titleWidth = 200),
  ),
  
  # Output: -----------------------------------------------------------------
  dashboardBody(
    # Define tabs per product
    tabsetPanel(
      id = "lic_data_set_both",
      tabPanel("ProductA", TS_module_UI("productA")),
      tabPanel("ProductB", TS_module_UI("productB")),
      tabPanel("ALL", TS_module_UI("all"))
      
    )
  )

server <- function(input, output) {
  x     <- reactive(input$year)
  y     <- reactive(input$y)
  prod_group <- reactive(input$prod_group)
  show_data <- reactive(input$show_data)
  
  callModule(
    TS_module,
    "productA",
    data = list(lic_data_set_both, mapdata, worldmapdata),
    product = "ProductA",
    prod_group,
    x,
    y,
    show_data
  )
  callModule(
    TS_module,
    "productB",
    data = list(lic_data_set_both, mapdata, worldmapdata),
    product = "ProductB",
    prod_group,
    x,
    y,
    show_data
  )
  callModule(
    TS_module,
    "all",
    data = list(lic_data_set_both, mapdata, worldmapdata),
    product = "ALL",
    prod_group,
    x,
    y,
    show_data
  )
}

shinyApp(ui = ui, server = server)

```

Here is an extract of the Module UI and server piece.
```{r, eval = FALSE}

# Module UI -------------------------------------------------------------------

TS_module_UI <- function(id) {
  ns <- NS(id)
  
  tagList(
   
    fluidRow(
      box(highchartOutput(ns("product_group"))),
      box(highchartOutput(ns("by_region"))),
      box(highchartOutput(ns("top_10_prod"))),
      box(highchartOutput(ns("top_10_states")))
    )
  )
}

#Module Server ----------------------------------------------------------------------------

TS_module <-
  function(input,
           output,
           session,
           data,
           product,
           prod_group,
           x,
           y,
           show_data) {
    
    TS_with_year_filtered <- reactive({
      
      data <- as_tibble(data[[1]])
      get_year <- x()
      get_unit_measure <- y()
      
      if (prod_group()) {
        data %>%
          filter(between(FY, get_year[1], get_year[2]),
                 `Prod Group` == "SERVER")
      } else {
        data %>% filter(between(FY, get_year[1], get_year[2]))
      }
    })
    
    
    TS_with_All_vs_namedProdtype <- reactive({
      
      if (as.character(product) == "ALL") {
        TS_with_year_filtered()
      } else {
        TS_with_year_filtered() %>%
          filter(Product == as.character(product))
      }
    })
    
 
    TS_with_Product <- reactive({
      get_unit_measure <- y()
      Product_data <- TS_with_year_filtered()
      
      
      if (as.character(product) == "ALL") {
        
        Product_data %>%
          group_by_at(.vars = c("FY", "SalesTeam")) %>%
          summarize_at(vars(get_unit_measure), sum)
      } else {
        Product_data %>%
          filter(Product == as.character(product)) %>%
          group_by_at(.vars = c("FY", "SalesTeam")) %>%
          summarize_at(vars(get_unit_measure), sum)
      }
    })
    
    
    #Calculate Total YTD data
    
    TS_YTD_Orders <- reactive({
      
      TS_with_Product() %>%
        summarise_at(.vars = c(y()), funs(Total = sum)) %>%
        summarise(Total = sum(Total)) %>%
        pull()
    })
    
    
    generate_top_10_states <- reactive({
      get_unit_measure <- y()
      TS_state_orders() %>%
        arrange_at(.vars = c(get_unit_measure), desc) %>%
        head(10)
      
    })
    
    output$top_10_states <- renderHighchart({
      
      df <- generate_top_10_states()
      get_unit_measure <- y()
      
      highchart() %>%
        hc_title(text = "Top 10: Orders_by_state") %>%
        hc_add_series(df,
                      "pie",
                      hcaes("`postal-code`",!!get_unit_measure),
                      name = as.character(get_unit_measure)) %>%
        hc_plotOptions(series = list(
          dataLabels = list(enabled = TRUE, format = '{point.name}: {point.percentage:.1f}%'))) %>%
        hc_exporting(
          enabled = T,
          buttons = list(contextButton = list(
            menuItems = c(
              "downloadPNG",
              "downloadJPEG",
              "separator",
              "downloadCSV",
              "downloadXLS"
            )
          )),
          filename = "Top_10_ordersbystate"
        )
    })
```

